DeepPicar: A Low-cost Deep Neural Network-based Autonomous Car

### 2018-01-30 paper/intro.tex
intro revision

We find that the DeepPicar is capable of 
completing control loops in under 33.$\overline{\mbox{33}}$ 
ms, or 30 hz, and can do so 100\% of the time.
Other tested embedded platforms, Intel UP and NVIDIA TX2, offer even
better performance, and are capable of supporting deep-learning based 
real-time control up to 100 Hz control frequency on the TX2 when the GPU 
is used. However, in all platforms, shared resource contention remains 
an important issue as we observe up to 11.6X control loop execution time
increase, mostly due to increase in the neural network inferencing
operation, when memory performance intensive applications are
co-scheduled on idle CPU cores.

% contributions
The {\bf contributions} of this paper are as follows:
\begin{itemize}
  \item We present the design and implementation of a
    low-cost autonomous vehicle testbed, DeepPicar, which utilizes
    state-of-the-art artificial intelligence techniques.
  \item We provide an analysis and case-study of real-time issues in the
    DeepPicar.
  \item We systematically compare real-time computing capabilities of
    multiple embedded computing platforms in the context of
    vision-based autonomous driving.
\end{itemize}

>>>>

% Findings:
%
% - we find DNN processing is an ideal real-time workload from the
%   perspective of timing predictability as the amount of computation
%   needed is fixed at design time and doesn't change over input.
%
% - real-time processing of non-trivial DNN inferencing is feasible
%   on today's embedded multicore platform, even as inexpensive as
%   raspberry pi3.
%
% - all tested platforms have enough computing capacity to consolidate 
%   multiple DeepPicar's CNN workloads or other compute intensive tasks
%   simultaneously.
%
% - however, when consolidating multiple workloads shared resource
%   contention is an important issue. we carefully analyzed
%   sensitivity of DeepPicar's workload with respect to shared
%   hardware resources.
%
% - in particular, we find DNN processing is highly sensitive to
%   memory performance. performance severely affected when memory
%   bandwidth is contented.
%
% - however, we find that dnn processing is not sensitive to cache
%   space. we find applying cache partitioning does not help improve
%   isolation of DNN workload when memory bandwidth is contended.
From the systematic study, we made a number of interesting
observations from the perspective of real-time systems.
% predictability
First, we find that DNN inferencing is, in a sense, an ideal real-time
workload as the amount of computation needed to compete a single
inference is fixed at the DNN architecture design time and does not
change at runtime over different input (e.g., different image frame).
This means that it is feasible to achieve highly predictable and
tightly bounded worst-case execution time estimation.
Second,

% performance on low-cost embedded multicore
Second, we find that real-time processing of DeepPicar's DNN-based
control loop is feasible on today's embedded computing platforms, even 
as inexpensive as Raspberry Pi 3. We find that the control loop completes 
under 33.$\overline{\mbox{33}}$ ms, or 30 Hz, using just two cores of
the Raspberry Pi 3's quad-core CPU (w/o using its GPU), and can
do so 100\% of the time; we can achieve 20 Hz control frequency using
just one core (out of four) of the Raspberry Pi 3's quad-core CPU.
Other tested embedded platforms, Intel UP and NVIDIA TX2, offer even
better performance, and are capable of supporting deep-learning based 
real-time control over 100 Hz control frequency on the TX2 when its
embedded GPU is used.
We find that all the tested embedded computing platforms have 
enough computing capacity to consolidate multiple instances of
DeepPicar's DNN workload.

% all platforms have enough capacity to consolidate
Third, contention in shared hardware resources remains an important
issue when consolidating multiple workloads. 
In particular, we find that DNN inferencing is highly sensitive to
memory performance inference as we observe up to 11.6X control loop
execution time increase, when memory performance intensive
applications are co-scheduled on idle CPU cores.
We find, however, that DNN inferencing is not sensitive to cache
space; we show that applying cache partitioning has negligible impact
in providing timing isolation on consolidated scenarios. This finding
suggests that in order to guarantee real-time performance of DNN-based
control applications, solutions to guarantee its memory performance
are much more important.

% contributions
The {\bf contributions} of this paper are as follows:
\begin{itemize}
  \item We provide the design and implementation of a low-cost
    autonomous car testbed, DeepPicar, which is based on a
    state-of-the-art end-to-end deep learning, as
    open-source~\footnote{The source code, build 
instruction, and a collection of self-driving videos of the DeepPicar
can be found at: \url{https://github.com/heechul/picar}}, with the
    goal of lowering the economic and safety-related barriers to study
    autonomous cars.

  \item We systematically analyze the characteristics of DeepPicar's
    DNN inferencing workload on a number of representative
    contemporary embedded computing platforms from the perspective of
    real-time systems. In particular, to the best of our knowledge,
    our work is first to investigate the impacts of shared resource
    contention on DNN inferencing workloads in the context of
    real-time systems.

\end{itemize}

### 2018-01-30 paper/related.tex
intro revision

\section{Related Work}
% discusstion
% - model is still quite small. cam we use bigger CNNs?
\fixme{What we want to say:
  1) We showed DNN is feasible on today's embedded computing
  platforms.
  2) But, the CNN model is rather smaller-kind with respect to today's
  state-of-the-art CNNs, which are much bigger.
  3) show table, borrowed from google's object detection api
  paper~\cite{huang2017speed}. 
  4) Applying these much more complex DNNs would improve network accuracy
  but they are obviously challenging for embedded computing
  platforms. We already discussed computing platform improvement that
  improve performance of executing a given DNN workload in
  section 2, so here we discuss model compression that reduces
  the necessary computing load itself.}

@@

% related work on testbeds
\fixme{discuss MIT racecar, UPENN car, donkey car and other testbeds?
  The points of discussion may include: 1) they are expensive. 2)
  donkey car is similar to us, but still a bit more expnsive, and more
  importantly there is no existing studies which systematically evaluate
  real-time performance of using DNN on embedded computing
  platforms. Our work provides quantitive real-time performance
  evaluation results, using a known DNN used in a real self-driving
  car. Our work is fully reproducible at a very low cost, lowring the
  barrier to entry in researching self-driving cars, helping the
  community blah blah blah.}

>>>>

\section{Related Work}
% discusstion
% - model is still quite small. cam we use bigger CNNs?
\fixme{What we want to say:
  1) We showed DNN is feasible on today's embedded computing
  platforms.
  2) But, the CNN model is rather smaller-kind with respect to today's
  state-of-the-art CNNs, which are much bigger.
  3) disscuss google's object detection api
  paper~\cite{huang2017speed}. 
  4) Applying these much more complex DNNs (such as google's detection
  models) may be challenging for embedded computing
  platforms.
  5) Two general approaches: (1) make it faster (computing workload
  remains the same) (2) reduce the workload itself (model compression
  or reduced precision etc.)}

@@

% related work on testbeds
\fixme{maybe we can also discuss MIT racecar, UPENN car, donkey car and other testbeds?
  The points of discussion may include: 1) they are expensive. 2)
  donkey car is similar to us, but still a bit more expnsive, and more
  importantly there is no existing studies which systematically evaluate
  real-time performance of using DNN on embedded computing
  platforms. Our work provides quantitive real-time performance
  evaluation results, using a known DNN used in a real self-driving
  car. Our work is fully reproducible at a very low cost, lowring the
  barrier to entry in researching self-driving cars, helping the
  community blah blah blah.}
